---
title: "Anomaly Detection"
author: "Faith Rotich"
date: "2/6/2022"
output: html_document
---

ANOMALY DETECTION

PROBLEM DEFINITION
a) Specifying the Question
Identify anomalies in the dataset = fraud detection

b) Defining the metrics for success
check whether there are any anomalies in the given sales dataset. The objective of this task being fraud detection.

c) Understanding the context
You are a Data analyst at Carrefour Kenya and are currently undertaking a project that will inform the marketing department on the most relevant marketing strategies that will result in the highest no. of sales (total price including tax). Your project has been divided into four parts where you’ll explore a recent marketing dataset by performing various unsupervised learning techniques and later providing recommendations based on your insights.

d) Recording the Experimental Design
Define the question, the metric for success, the context, experimental design taken.
Read and explore the given dataset.
Identify anomalies in the dataset

e) Relevance of the data
You have also been requested to check whether there are any anomalies in the given sales dataset. The objective of this task being fraud detection


Dataset link - ["http://bit.ly/CarreFourSalesDataset"]




```{R}
# Installing anomalize package
# ---
#install.packages("anomalize")
```


# 1. Loading the required packages.


```{r}
library(anomalize) 
library(tidyverse) 
library(tibble)
library(tibbletime) 
library(data.table)
library(dplyr)
```


# 2. Loading and Viewing the dataset

```{r}
data<- read.csv("http://bit.ly/CarreFourSalesDataset")

```


```{r}
#Viewing the first 5 rows 

head(data) 

```
The dataset contains 2 columns
The Sales column shows values of sales and particular dates shows the dates of sales


```{r}
# Viewing the last 5 row s
tail(data)
```


Checking the dimensions of the dataset.
```{r}
# Checking the size of the dataset
dim(data)
```
It has 1000 observations on two variables that is dates and sales 


# 3. Data Cleaning 

Checking for missing values.

```{r}
colSums(is.na(data))
```
There are no missing values in the dataset



Checking the datatypes
```{r}
#data structure
str(data)
```
Date column is a string and it needs to be converted to date using as.Date()
 
 
 
 
```{r}
data$Date <- as.Date(data$Date, format = "%m/%d/%Y")
data$Date <- sort(data$Date, decreasing = FALSE)
```





#Rechecking the structure to confirm the operation 
```{r}
#data structure
str(data)
```

# 4. Feature Engineering

```{r}
### Converting the sales into a tibbletime object

library(timetk)

```



```{r}
data$Date <- as.POSIXct(data$Date)
```


# Loading libraries for anomaly detection
```{r}
library(tibble)
library(tibbletime) 
```


# Loading the ggplot2 library for plotting 
```{r}
library(ggplot2)
```




```{r}
# Using as_tibble function
data <- as_tibble(data)
```




# 5.Anomaly Detection.



```{r}
# Decomposing 

data %>%
time_decompose(Sales, method = 'stl', frequency = 'auto', trend = 'auto') %>%
anomalize(remainder, method = 'gesd', alpha = 0.1, max_anoms = 0.5) %>%
plot_anomaly_decomposition(ncol = 3, alpha_dots = 0.7)
```



```{r}

# Recomposing 
# Using time_recompose() function to generate bands around the normal levels of observed values for clear observation of the outliers 

data %>%
time_decompose(Sales, method = 'stl', frequency = 'auto', trend = 'auto') %>%
anomalize(remainder, method = 'gesd', alpha = 0.1, max_anoms = 0.1) %>%
time_recompose() %>%
plot_anomalies(time_recomposed = TRUE, ncol = 3, alpha_dots = 0.5)


```


Checking the time periods that had anomalies

```{r}
anomalies = data %>%
time_decompose(Sales, method = 'stl', frequency = 'auto', trend = 'auto') %>%
anomalize(remainder, method = 'gesd', alpha = 0.05, max_anoms = 0.1) %>%
time_recompose() %>%
filter(anomaly == 'Yes')
anomalies
```

# 6. Conclusion
There are several anomalies in the Carrefour dataset  between the 6th of February 2019 and 30th of March 2019 as shown in the table above


# 7. Recommendations
Carrerfour should review the sales patterns between February and March  and investigate the reason why there were why anomalies between these two months.


